---
title: "Lydia Myers ADA HW 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Challenge 1


Loading data into R
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
```

```{r}

f<-"https://raw.githubusercontent.com/difiore/ada-2021-datasets/main/darwin.txt"
darwin<-read.delim(f, header=T, sep="\t")
```

How many paragraphs are there in the chapter? 

```{r}
glimpse(darwin)
```
There are 56. 

Printing the 34th paragraph
```{r}
x<-darwin[34,]
print(x)
```
Splitting the 34th paragraph into a vector of 1-word character strings
```{r}
split<-str_split(x, " ")
split<-unlist(split)
```
Looking into the structure of the new split vector
```{r}

unique<-unique(split)
length(unique)

names(table(split))[as.vector(table(split))==max(table(split))]

frequencies<-table(split)
length(names(frequencies)[frequencies==1])

length(names(frequencies)[frequencies>4])
```
There are 129 unique words in the vector, the most common of which is 'of'. 98 words appear only once, and 5 words appear 5 or more times.

Extracting the text of the last paragraph
```{r}
final_quote<-darwin[56,]
```
Splitting the last paragraph
```{r}
split_quote<-str_split(final_quote, " ")
split_quote<-unlist(split_quote)
```
Extracting every 3rd word of the last paragraph
```{r}
every_third<-split_quote[seq(1,length(split_quote), 5)]
print(every_third)
```
Arranging the resulting vector into reverse alphabetical order
```{r}
sort(every_third, decreasing=T)
```

## Challenge 2


Creating the vectors
```{r}
t<-c(35,88,42,84,81,30)
city<-c("Beijing","Lagos","Paris","Rio de Janeiro","San Juan","Toronto")
```

Associating the temperature data with the city data
```{r}
names(t)<-city
```
Accessing the temperature of the first three cities on the list
```{r}
t[1:3]
```
Accessing the temperatures of Paris and San Juan
```{r}
t[c(3,5)]
```

## Challenge 3

Creating the matrix
```{r}
m1<-matrix(data=c(159:1),nrow=8, ncol=20)
```
Extracting the number at row 5, column 2
```{r}
m1[5,2]
```
Extracting rows 5-7
```{r}
m1[c(5:7),]
```
Extracting the contents of row3 column4 to row6 column9 as a new variable
```{r}
m2<-m1[3:6,4:9]
m2
```
Examining this variable
```{r}
class(m2)
mode(m2)
```
This variable is a matrix array, and the mode is numeric.

## Challenge 4

Creating the array
```{r}
a<-array(data=seq(2,800,2), dim=c(5,5,4,4))
```
Testing to see what the examples return
```{r}
a[1,1,1,2]
a[2,3,2,]
a[1:5,1:5,3,3]
```

## Challenge 5

Creating the list
```{r}
#strepsirrhines
lorisoidea<-c("Lorisidae","Galagidae")
lorisiformes<-list(lorisoidea)
names(lorisiformes)<c("Lorisoidea")
lemuroidea<-c("Cheirogaleidae","Lepilemuridae","Indriidae","Lemuridae","Daubentoniidae")
lemuriformes<-list(lemuroidea)
names(lemuriformes)<-c("Lemuroidea")
strepsirrhini<-list(lorisiformes, lemuriformes)
names(strepsirrhini)<-c("Lorisiformes","Lemuriformes")

#haplorhines
cercopithecoidea<-c("Cercopithecidae")
hominoidea<-c("Hylobatidae","Hominidae")
catarhini<-list(hominoidea, cercopithecoidea)
names(catarhini)<-c("Hominoidea","Cercopithecoidea")
ceboidea<-c("Cebidae","Atelidae","Pitheciidae")
platyrrhini<-list(ceboidea)
names(platyrrhini)<-c("Ceboidea")
simiiformes<-list(platyrrhini, catarhini)
names(simiiformes)<-c("Platyrrhini", "Catarhini")
tarsioidea<-c("Tarsiidae")
tarsiiformes<-list(tarsioidea)
names(tarsiiformes)<-c("Tarsioidea")
haplorhini<-list(tarsiiformes, simiiformes)
names(haplorhini)<-c("Tarsiiformes","Simiiformes")
primates<-list(strepsirrhini, haplorhini)
names(primates)<-c("Strepsirrhini","Haplorhini")

#all primates
all<-list(primates)
names(all)<-c("Primates")
```
Separating platyrrhines
```{r}
nwm<-haplorhini[[2]][1]
```
Looking at the structure of this new variable
```{r}
class(nwm)
mode(nwm)
```
'nwm' is a list, and the mode is also list.

Extracting tarsiers
```{r}
haplorhini$Tarsiiformes
```
